The impacts of park access on health outcomes: a spatial comprehensive approach

### 2019-01-07 manuscript.Rnw
Restructure data tables in paper

@@

\begin{table*}
  \caption{Summary Statistics of Tract Data}
  \label{tab:tractsdata}
<<data_summary, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
our_summary <- list(
  "obesity"     = tab_summary(tracts$obesity),
  "physact"     = tab_summary(tracts$physact),
  "pop_density" = tab_summary(tracts$density),
  "fulltime"    = tab_summary(tracts$fulltime),
  "college"     = tab_summary(tracts$college)
)

summary_table(tracts@data, our_summary) %>%
  kable("latex", booktabs = T, escape = FALSE) %>%
  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
  group_rows("Physical Activity: percent of tract reporting physical activity", 5, 8) %>%
  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
  group_rows("College: percent of adults with a college degree", 17, 20)
@
\end{table*}

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance. Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space.

@@

\begin{table*}
  \caption{Summary Statistics of Parks Data}
  \label{tab:parksdata}
<<parks_summary, results="asis", message=FALSE, warning=FALSE, cache = FALSE>>=
park_summary <- list(
  "size"     = tab_summary(parks$size),
  "tweets"   = tab_summary(parks$tweets)
)

summary_table(as_data_frame(parks), park_summary) %>%
  kable("latex", booktabs = T, escape = FALSE, col.names = c("Parks"))
@
\end{table*}

@@

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with a semi-log specification:
\begin{equation}\label{eq:themodel}
\log(y_i) = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
\end{equation}
<<base_formula>>=
base_formula <- formula(
  ~ log(density) +
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)
@
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$CS_i$ is a choice-based
accessibility measure for the tract computed from Equation \ref{eq:cs} and
$g(W, \vect{y}, X)$
represents the impact of spatial neighborhood effects on the outcome.

@@

<<weights>>=
pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(2263)

@@

\begin{table*}
  \caption{Estimated Coefficients Relating Park Access to Physical Activity}
  \label{tab:pa_models}
<<pamodelchart, results="asis">>=
pa_models <- list(Base = physact_base_sdm, `Access` = pa_access_sdm,
                       `Access + Tweets` = pa_tweets_sdm)
texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
       booktabs = TRUE, use.packages = FALSE)
#' screenreg(pa_models, digits = 4, table = FALSE)
@
\end{table*}

@@

\begin{table*}
  \caption{Simulated Impact of Accessibility on Physical Activity}
  \label{tab:pa_impacts}
<<impacts>>=
physact_impacts <- pa_models %>%
  lapply(extract_impacts) %>%
  bind_rows(.id = "model") %>% ungroup() %>%
  rename(submodel = effect)

@@

physact_impacts %>%
  transmute(
    model, term, submodel, estimate,
    output = paste(round(`p-val`, 5),
                   gtools::stars.pval(`p-val`))
  ) %>%
  filter(term %in% c("access_ls", "tweets_ls")) %>%
  select(-term) %>%
  kable("latex", booktabs = T, escape = FALSE,
        col.names = c("Logsum Measure", "Effect", "Estimate", "$p$-value"))
@
\end{table*}

@@

It is of course necessary to also consider the magnitude of the effect.
As this is a semi-log model, the coefficient implies a unit increase
in the access and tweets logsum leads to a
\Sexpr{round(effect_estimate * 100, 3)}\% increase in the physical activity rate.
Referring to the map of logsum measures in Figure \ref{fig:map}, the measures
span from roughly $-3$ in heavily industrialized segments of southern Brooklyn
to roughly $3$ in tracts immediately adjacent to Central Park on the Upper East
Side. Thus moving from the least accessible parts of the city to the most
accessible will reduce the expected physical activity rate by
\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. For an aggregate statistic, this
is not an insubstantial marginal effect.

<<obesity_models>>=
obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ . + log_physact),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
obese_access_sdm <- update(obese_base_sdm, . ~ . + access_ls + access_ls * log_physact)
obese_tweets_sdm <- update(obese_base_sdm, . ~ . + tweets_ls + access_ls * log_physact)
obesity_models <- list("Obesity" = obese_base_sdm, "Obesity Access" = obese_access_sdm, 
                       "Obesity Tweets + Access" = obese_tweets_sdm)
@

<<obese_impacts>>=
obese_impacts <- obesity_models %>%
  lapply(extract_impacts) %>%
  bind_rows(.id = "model") %>% ungroup() %>%
  rename(submodel = effect)

dwplot(bind_rows(physact_impacts, obese_impacts) %>% 
                 filter(term %in% c("log_physact", "access_ls", "tweets_ls")) %>%
                 filter(submodel == "Total")) +
  theme_bw()

obese_impacts %>%
  transmute(
    model, term, submodel, estimate,
    output = paste(round(`p-val`, 5),
                   gtools::stars.pval(`p-val`))
  ) %>%
  filter(term %in% c("log_physact", "access_ls", "tweets_ls"), submodel == "Total")
@

@@

>>>>

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",", drop0trailing = TRUE)
})
@

@@

@@

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance. Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space. This 
leaves us with \Sexpr{nrow(parks)} separate parks.

<<tractsdata, results="asis", message=FALSE, warning=FALSE, cache=FALSE>>=
tracts_data_index <- data_frame(
  variable =  c(
    "obesity", "physact", "density", "fulltime", "college", "single", 
    "Pct0to17", "Pct18to29",  "Pct65plus", "black", "asian", "hispanic", 
    "other")
  ) %>% mutate(order = row_number())

tract_summary <- tracts@data %>%
  dplyr::select(obesity:Pct65plus) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  inner_join(tracts_data_index) %>%
  arrange(order) %>%
  mutate(
    variable = c(
      "Obesity Rate", "Physical Activity Rate",
      "Density: population per square mile",
      "Full-time workers", "College-educated", 
      "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)", 
      "Share of seniors (65+)", "Black population share", "Asian population share", 
      "Hispanic population share", "Other Minorities")
  ) %>%
  dplyr::select(-order)

park_summary <- parks %>%
  st_set_geometry(NULL) %>%
  dplyr::select(size, tweets) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  mutate( variable = c("Park Size in Acres", "Tweets Emanating from Park"))

bind_rows(tract_summary, park_summary) %>%
  kable(escape = FALSE, digits = 2, 
        table.envir = "table*", booktabs = TRUE,
        caption = "Descriptive Statistics of Tract and Park Variables",
        col.names = c(" ", "Minimum", "Median (IQR)", "Maximum")) %>%
  kable_styling() %>%
  group_rows(paste("Tract Variables, N =", nrow(tracts@data)), 1, 13) %>%
  group_rows(paste("Park Variables, N =", nrow(parks)), 14, 15)

@@

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with a semi-log specification:
\begin{equation}\label{eq:themodel}
\log(y_i) = f(\vect{x}_i, CS_i, g(W, \mathit{Obesity}, \vect{X}))
\end{equation}
<<base_formula>>=
base_formula <- formula(
  ~ log(density) +
    fulltime + college + single +
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity
    black + asian + hispanic + other)

termnames <- c("log(density)" , "Full-time workers", "College-educated", 
               "Single Adults", "Share of children (0-17)", "Share of young adults (18-29)", 
               "Share of seniors (65+)", "Black population share", "Asian population share", 
               "Hispanic population share", "Other Minorities")
@
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$CS_i$ is a choice-based
accessibility measure for the tract computed from Equation \ref{eq:cs} and
$g(W, \vect{y}, X)$
represents the impact of spatial neighborhood effects on the outcome.

@@

<<weights, warning = FALSE>>=
pop_weighted_centroids <- read_csv("data/CenPop2010_Mean_TR36.csv") %>%
  mutate(geoid = str_c(STATEFP, COUNTYFP, TRACTCE))
pop_weighted_centroids <- left_join(tracts@data, pop_weighted_centroids, by = "geoid") %>%
  dplyr::select(geoid, LONGITUDE, LATITUDE) %>%
  st_as_sf(crs = 4326, coords = c("LONGITUDE", "LATITUDE")) %>%
  st_transform(2263)

@@

\begin{table*}
  \caption{Estimated Coefficients Relating Park Access to Physical Activity}
  \label{tab:pa_models}
<<pamodelchart, results="asis">>=
pa_models <- list(Base = physact_base_sdm, `Access` = pa_access_sdm,
                       `Access + Tweets` = pa_tweets_sdm)
texreg(pa_models, digits = 4, table = FALSE, single.row = TRUE,
       booktabs = TRUE, use.packages = FALSE, 
       custom.coef.names = c("Intercept", termnames, paste("lag", termnames), "$\\rho$",
                             "Access logsum", "lag Access logsum", 
                             "Access + Tweets logsum", "lag Access + Tweets logsum"))
#' screenreg(pa_models, digits = 4, table = FALSE)
@
\end{table*}

@@

\begin{table*}
  \caption{Simulated Impact of Accessibility on Physical Activity}
  \label{tab:pa_impacts}
<<impacts, results="asis">>=
physact_impacts <- pa_models %>%
  lapply(extract_impacts) %>%
  bind_rows(.id = "model") %>% ungroup() %>%
  rename(submodel = effect)

@@

physact_impacts %>%
  transmute(
    model, term, submodel, estimate,
    output = paste(round(`p-val`, 5),
                   gtools::stars.pval(`p-val`))
  ) %>%
  filter(term %in% c("access_ls", "tweets_ls"), submodel == "Total") %>%
  dplyr::select(-term, -submodel) %>%
  kable("latex", booktabs = T, escape = FALSE,
        col.names = c("Logsum Measure", "Estimate", "$p$-value")) %>%
  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
@
\end{table*}

@@

It is of course necessary to also consider the magnitude of the effect.
As this is a semi-log model, the coefficient implies a unit increase
in the access and tweets logsum leads to a
\Sexpr{round(effect_estimate * 100, 3)}\% increase in the physical activity rate.
Referring to the map of logsum measures in Figure \ref{fig:map}, the measures
span from roughly $-3$ in heavily industrialized segments of southern Brooklyn
to roughly $3$ in tracts immediately adjacent to Central Park on the Upper East
Side. Thus moving from the least accessible parts of the city to the most
accessible will reduce the expected physical activity rate by
\Sexpr{round(effect_estimate * 100 * 6, 3)}\%. In other words, if a tract with 
poor accessibility is expected to achieve the sample mean of 
\Sexpr{round(mean(tracts$physact), 1)}\%, the same tract with excellent accessibility
is expected to achieve a rate of 
\Sexpr{round(mean(tracts$physact) * (1 + effect_estimate  * 6), 2)}\%. 
For an aggregate statistic strongly influenced by other socioeconomic factors, 
this is not an insubstantial marginal effect.

<<obesity_models>>=
obese_tweets_sdm <- update(pa_tweets_sdm, log_obesity ~ . + log_physact)
@

We now consider the impacts of a model where the dependent variable is 
the {\em obesity} rate, and physical activity becomes an independent covariate
alongside the controlling variables and the tweets-augmented logsum variable. 
Table \ref{tab:obese_impacts} presents the simulated impacts resulting from
this model.

\begin{table*}
  \caption{Simulated Impacts of Variables on Obesity}
  \label{tab:obese_impacts}
<<tab:obese_impacts, results="asis">>=
extract_impacts(obese_tweets_sdm) %>%
  group_by(effect) %>%
  transmute(
    term,
    output = paste(round(`estimate`, 5),
                   gtools::stars.pval(`p-val`)),
    order = row_number()
  ) %>%
  spread(effect, output) %>%
  arrange(order) %>%
  dplyr::select(-order) %>%
  mutate(
    term = c(termnames, "Access + Tweets logsum", "log(Physical Activity Rate)")
  ) %>%
  kable("latex", booktabs = T, escape = FALSE, digits = 3, 
        col.names = c("", "Direct", "Indirect", "Total")) %>%
  footnote("$^{***}p<0.001$, $^{**}p<0.01$, $^*p<0.05$", escape = F)
@
\end{table*}

@@

We readily acknowledge limitations in this study. To begin, it would be beneficial 
to have estimated the logsum coefficients from a high-quality survey rather than
asserting them to match {\em a priori expectations of}
